### 2020-02-18 R/supp_material.Rmd
added supp mat document and edited theoretical results figure

---
bibliography: Whitebib.bib
csl: ecology-letters.csl
editor_options:
  chunk_output_type: console
fontsize: 12pt
geometry: margin=1in
header-includes: \usepackage{float} \usepackage{lineno} \usepackage{setspace}\doublespacing
  \usepackage[round]{natbib} \bibpunct[; ]{(}{)}{,}{a}{}{,} \usepackage{color} \usepackage{totcount}
  \newtotcounter{citenum} \def\oldcite{} \let\oldcite=\bibcite \def\bibcite{\stepcounter{citenum}\oldcite}
  \renewcommand{\thepage}{S\arabic{page}}
  \renewcommand{\thesection}{Appendix S\arabic{section}} \renewcommand{\thetable}{S\arabic{table}}
  \renewcommand{\thefigure}{S\arabic{figure}}
  \usepackage{amsmath}
  \renewcommand{\theequation}{S\arabic{equation}}
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = dirname(getwd()))
```

\begin{center}
\textbf{\Large Supplementary Material for: Title here}
\vspace{7 mm}

\textsc{Rosalie Bruel$^{1}$ and Easton R. White$^{2}$}
\vspace{5 mm}

\normalsize{\indent $^{1}$Rubenstein Ecosystem Science Laboratory, University of Vermont, VT 05405, USA \\ $^{2}$Department of Biology, University of Vermont, VT 05405, USA}
\end{center}

\vspace{2 mm}

\tableofcontents

\vspace{1cm}

Data and code for all the figures can be found at (https://github.com/rosalieb/temporal-sampling.git). All the analyses were run in R.

\clearpage

# Additional figures

## Simulation approach

```{r theoretical_results,message=F,warning=F,echo=F,eval=F}

source('R/sampling_regular.R')
source('R/sampling_random.R')
source('R/build_time_series.R')
source('R/updated_param_sensitivity.R')

require(tidyverse)

par(mfrow=c(2,2))
# Sigma plot
sigma15 <- param_sensitivity('sigma',15)
plot(sigma15,ylab='Probability correct',xlab='sigma',pch=16,cex.lab=1.3,las=1,ylim=c(0,1),col=1)

sigma20 <- param_sensitivity('sigma',20)
points(sigma20,ylab='Probability correct',xlab='sigma',pch=15,cex.lab=1.3,las=1,ylim=c(0,1),col=3)

# phi plot
phi15 <- param_sensitivity('phi',15)
plot(phi15,ylab='Probability correct',xlab='phi',pch=16,cex.lab=1.3,las=1,ylim=c(0,1),col=1)

phi20 <- param_sensitivity('phi',20)
points(phi20,ylab='Probability correct',xlab='phi',pch=15,cex.lab=1.3,las=1,ylim=c(0,1),col=3)

```

```{r theoretical_results_min_time,message=F,warning=F,echo=F,eval=F}

source('R/sampling_regular.R')
source('R/sampling_random.R')
source('R/build_time_series.R')
source('R/updated_param_sensitivity.R')
source('R/calc_min_time2.R')
require(tidyverse)

# Sigma results
sigma_vec = seq(0.1,30,4)
min_time  = rep(0,times=length(sigma_vec))
sigma_output <- data.frame(sigma_vec,min_time)

for (param_num in 1:nrow(sigma_output)){
  sigma_output$min_time[param_num] <- calc_min_time('sigma',sigma_output$sigma_vec[param_num])
  print(sigma_output)
}

write.csv(x = sigma_output,file = 'Output/theoretical_results_sigma.csv',quote=FALSE)

# Phi results

phi_vec = seq(-0.8,0.8,0.2)
min_time  = rep(0,times=length(phi_vec))
phi_output <- data.frame(phi_vec,min_time)

for (param_num in 1:nrow(phi_output)){
  phi_output$min_time[param_num] <- calc_min_time('phi',phi_output$phi_vec[param_num])
  print(phi_output)
}

write.csv(x = phi_output,file = 'Output/theoretical_results_phi.csv',quote=FALSE)

# Shift size

shift_size_vec = rev(seq(0.1,0.9,0.025))
min_time  = rep(0,times=length(shift_size_vec))
shift_size_output <- data.frame(shift_size_vec,min_time)

for (param_num in 1:nrow(shift_size_output)){
  shift_size_output$min_time[param_num] <- calc_min_time('shift_size',shift_size_output$shift_size_vec[param_num])
  print(shift_size_output)
}

write.csv(x = shift_size_output,file = 'Output/theoretical_results_shift_size.csv',quote=FALSE)

```

```{r theoretical_results_min_time_plot, fig.cap='Minimum number of samples required for 0.8 statistical power given different levels of (a) temporal variability ($\\sigma$), (b) temporal autocorrelation ($\\phi$), and (c) shift size. The default parameters are $\\sigma = 3$, $\\phi = 0.5$, and a shift size $=0.75$. The exact timing of the true changepoint varied for each simulation between time steps 30 and 70. \\label{fig:theoretical_min_time}',message=F,warning=F,echo=F,eval=T}

par(mfrow=c(1,3))

sigma <- read.csv('Output/theoretical_results_sigma.csv',header=TRUE)
plot(sigma$sigma_vec,sigma$min_time,ylab='Minimum number of samples required',xlab='sigma',pch=16,cex.lab=1.3,las=1,ylim=c(20,50),col=1,type='o')

# Phi results

phi <- read.csv('Output/theoretical_results_phi.csv',header=TRUE)
plot(phi$phi_vec,phi$min_time,ylab=' ',xlab='phi',pch=16,cex.lab=1.3,las=1,ylim=c(20,50),col=1,type='o')

# Shift size results

shift_size <- read.csv('Output/theoretical_results_shift_size.csv',header=TRUE)
plot(shift_size$shift_size_vec,shift_size$min_time,ylab=' ',xlab='shift size',pch=16,cex.lab=1.3,las=1,ylim=c(20,50),col=1,type='o')

```

```{r theoretical_results_sampling_approaches,message=F,warning=F,echo=F,eval=F}

source('R/sampling_regular.R')
source('R/sampling_random.R')
source('R/build_time_series.R')
source('R/updated_param_sensitivity.R')
source('R/calc_min_time2.R')
require(tidyverse)

# Need to run each method on a set of simulated time series. We would then have 100 runs for each method and each sampling amount

approach <- c('random','regular','iterative')
number_samples <- 1:30
simulation_number <- 1:100
output_theory_approach <- data.frame(expand.grid(approach,number_samples,simulation_number))
names(output_theory_approach) <- c('approach', 'number_samples', 'simulation_number')
output_theory_approach$distance_to_real = 0

# Build function that takes in approach, simulated time series, and number of samples. The function should then spit out the distance to the real changepoint

# Set of time series (need other params)
params <- tibble(shift_time = sample(30:70,nrow(params),replace = T))

time_series <- params %>%
  pmap(build_time_series)

# Run through params

for (i in 1:nrow(output_theory_approach)){

  output_theory_approach$distance_to_real[i] <- calc_distance_to_real(output_theory_approach$approach,output_theory_approach$number_samples,time_series[output_theory_approach$simulation_number])
}

```

## Case study

```{r case study read data output, include=FALSE}
myoutput <- read.table(paste0(getwd(),"/Output/output_changepoint_varese.txt"))
myoutput$method_f <- factor(myoutput$method, levels=c('Random', 'Regular', 'Iterative'))

```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 7.2, fig.height= 8}
library(ggplot2)
library(gridExtra)
p1 <- ggplot(myoutput[myoutput$target_cpt==1,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="a. First changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
p2 <- ggplot(myoutput[myoutput$target_cpt==2,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="b. Second changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
p3 <- ggplot(myoutput[myoutput$target_cpt==3,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="c. Third changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
grid.arrange(p1,p2,p3,nrow=3)
```

__Main point__: Changepoint was detected even though we used a transformed vector (from community to first component of DCA)

## References