### 2020-03-30 R/supp_material.Rmd

---
bibliography: Whitebib.bib
csl: ecology-letters.csl
editor_options:
  chunk_output_type: console
fontsize: 12pt
geometry: margin=1in
header-includes: \usepackage{float} \usepackage{lineno} \usepackage{setspace}\doublespacing
  \usepackage[round]{natbib} \bibpunct[; ]{(}{)}{,}{a}{}{,} \usepackage{color} \usepackage{totcount}
  \newtotcounter{citenum} \def\oldcite{} \let\oldcite=\bibcite \def\bibcite{\stepcounter{citenum}\oldcite}
  \renewcommand{\thepage}{S\arabic{page}}
  \renewcommand{\thesection}{Appendix S\arabic{section}} \renewcommand{\thetable}{S\arabic{table}}
  \renewcommand{\thefigure}{S\arabic{figure}}
  \usepackage{amsmath}
  \renewcommand{\theequation}{S\arabic{equation}} \floatplacement{figure}{H}
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = dirname(getwd()))
```

\begin{center}
\textbf{\Large Supplementary Material for: Sampling requirements and approaches to detect ecosystem shifts}
\vspace{7 mm}

\textsc{Rosalie Bruel$^{1}$ and Easton R. White$^{2}$}
\vspace{5 mm}

\normalsize{\indent $^{1}$Rubenstein Ecosystem Science Laboratory, University of Vermont, VT 05405, USA \\ $^{2}$Department of Biology, University of Vermont, VT 05405, USA}
\end{center}

\vspace{2 mm}

\tableofcontents

\vspace{1cm}

Data and code for all the figures can be found at (https://github.com/rosalieb/temporal-sampling.git). All the analyses were run in R (citation).

\clearpage

# Additional text

For each sampling approach, we wrote three custom functions:

- sample_random(): sample first and last sample, in addition of n-2 random samples in between (n = maximum number of sample, chosen by the user),  
- sample_regular(): sample first and last sample, in addition of n-2 evenly spaced samples in between (n = maximum number of samples, chosen by the user), and  
- sample_iterative(): sample first and last sample as well as 3 evenly spaced samples in between (3 is the default but can be modified), in order to initiate the changepoint detection. Upon detection of the changepoint on this 5 samples time series, a new sample is added between the changepoint and the previous sample, to narrow down the real changepoint. Detection of changepoint and addition of sample is repeated, until it was narrowed down to two consecutive samples, or n (maximum number of samples, chosen by the user) was reached, whichever comes first.

# Parameter sensitivity

```{r parameter_sensitivity,message=F,warning=F,echo=F,eval=F,cache=F}
# Build 1x3 plot with sigma vs phi with color equal to distance to changepoint, power, or min time. The three panel plots would represent different shift size values

source('R/sampling_regular.R')
source('R/sampling_random.R')
source('R/build_time_series.R')
source('R/calc_min_time2.R')
require(tidyverse)

number_sims <- 10
# Find minimum time for different parameter combos

sigma_vec = seq(0.01,0.3,0.02)
phi_vec = seq(-0.8,0.8,0.1)
shift_size_vec = c(0.1,0.4,0.8)

output_theory_approach <- data.frame(expand.grid(sigma_vec,phi_vec,shift_size_vec))
names(output_theory_approach) <- c('sigma_vec', 'phi_vec', 'shift_size_vec')
output_theory_approach$power = 0

#params <- tibble(shift_time = sample(30:70,max(simulation_number),replace = T))

shift_time = sample(30:70,number_sims,replace=T) # Keep shift times constant for each param combo

# Run through params
for (i in 1:nrow(output_theory_approach)){

# Build XX number of time series
params <- list(sigma=rep(output_theory_approach$sigma_vec[i],number_sims),phi=rep(output_theory_approach$phi_vec[i],number_sims),shift_size=rep(output_theory_approach$shift_size_vec[i],number_sims),shift_time = shift_time)

time_series <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) sample_regular(df,20)$changepoint)

power <- sum(abs(time_series-shift_time)<5)/number_sims

#print(i)
output_theory_approach$power[i] <- power
}

write.csv(x = output_theory_approach,file = 'Output/parameter_sensitivity.csv',quote = FALSE )

require(ggplot2)
require(viridis)
ggplot(aes(x=phi_vec,y=sigma_vec,fill=power),data=output_theory_approach) + geom_tile() +
   scale_fill_viridis(discrete=FALSE,begin = 0.1,end=0.9) +
  facet_wrap(~shift_size_vec) +
   labs(fill='Statistical power') +
   xlab("Lag-1 autocorrelation (phi)") +
   ylab("Standard deviation (sigma)") +
   theme_classic(base_size = 14, base_family = "")

```

```{r parameter_sensitivity_plot,echo=F,message=F,warnings=F,fig.height=4,fig.cap='Regular sampling statistical power (fraction of 100 simulations which detecting a changepoint within five time points of the true changepoint) for different levels of standard deviation, lag-1 autocorrelation, and shift size.\\label{fig:parameter_sensitivity_plot}'}
output_theory_approach <- read.csv(file = 'Output/parameter_sensitivity.csv')

require(ggplot2)
require(viridis)
ggplot(aes(x=phi_vec,y=sigma_vec,fill=power),data=output_theory_approach) + geom_tile() +
   scale_fill_viridis(discrete=FALSE,begin = 0.1,end=0.9) +
  facet_wrap(~shift_size_vec) + 
   theme_classic(base_size = 14, base_family = "")+
  theme(plot.subtitle = element_text(hjust = 0.5)) +
   labs(fill='Power',subtitle = 'Shift size') +
   xlab("Lag-1 autocorrelation (phi)") +
   ylab("Standard deviation (sigma)")
```

# Minimum time for other sampling approaches

## Random sampling

```{r simulation_min_time_random,message=F,warning=F,echo=F,eval=F,cache=T}

source('R/sampling_random.R')
source('R/sampling_regular.R')
source('R/sampling_iterative.R')
source('R/build_time_series.R')
source('R/updated_param_sensitivity.R')
source('R/calc_min_time2.R')
require(tidyverse)

# Sigma results
sigma_vec = seq(0.01,0.4,0.025)
min_time  = rep(0,times=length(sigma_vec))
sigma_output <- data.frame(sigma_vec,min_time)

for (param_num in 1:nrow(sigma_output)){
  sigma_output$min_time[param_num] <- calc_min_time('sigma',sigma_output$sigma_vec[param_num],'random')
  #print(sigma_output)
}

write.csv(x = sigma_output,file = 'Output/theoretical_results_random_sigma.csv',quote=FALSE)

# Phi results

phi_vec = seq(-0.8,0.8,0.05)
min_time  = rep(0,times=length(phi_vec))
phi_output <- data.frame(phi_vec,min_time)

for (param_num in 1:nrow(phi_output)){
  phi_output$min_time[param_num] <- calc_min_time('phi',phi_output$phi_vec[param_num],'random')
  #print(phi_output)
}

write.csv(x = phi_output,file = 'Output/theoretical_results_random_phi.csv',quote=FALSE)

# Shift size

shift_size_vec = rev(seq(0.1,0.9,0.05))
min_time  = rep(0,times=length(shift_size_vec))
shift_size_output <- data.frame(shift_size_vec,min_time)

for (param_num in 1:nrow(shift_size_output)){
  shift_size_output$min_time[param_num] <- calc_min_time('shift_size',shift_size_output$shift_size_vec[param_num],'random')
  #print(shift_size_output)
}

write.csv(x = shift_size_output,file = 'Output/theoretical_results_random_shift_size.csv',quote=FALSE)

```

```{r simulation_min_time_plot_random, fig.width = 7.2, fig.height= 3, fig.cap='Minimum number of samples required for 0.8 statistical power given different levels of (a) temporal variability ($\\sigma$), (b) temporal autocorrelation ($\\phi$), and (c) shift size. Random sampling was used with the default parameters are $\\sigma = 0.53$, $\\phi = 0.404$, and a shift size $=0.81$ to match the case study. The exact timing of the true changepoint varied for each simulation between time steps 30 and 70. Each vertical line indicates the respective parameter calculated from the case study time series. \\label{fig:simulation_min_time_random}',message=F,warning=F,echo=F,eval=T}
require(ggplot2)
require(dplyr)
require(stickylabeller)
par(mfrow=c(1,3))

sigma <- read.csv('Output/theoretical_results_random_sigma.csv',header=TRUE)

phi <- read.csv('Output/theoretical_results_random_phi.csv',header=TRUE)

shift_size <- read.csv('Output/theoretical_results_random_shift_size.csv',header=TRUE)

# Melt the 3 dataset to be able to use facet_wrap()
colnames(sigma) <- colnames(phi) <- colnames(shift_size) <- c("X", "Param", "min_time")
simulation_all <- 
  rbind(cbind("Parameter" = rep("Standard deviation", nrow(sigma)), sigma),
        cbind("Parameter" = rep("Autocorrelation at lag-1", nrow(phi)), phi),
        cbind("Parameter" = rep("Normalized amplitude of shift", nrow(shift_size)), shift_size))
# param_vr <- data.frame("Parameter" = unique(simulation_all$Parameter),
#                        "value" = c(ar_vr, cv_vr, shift_vr))

p1 <- 
  ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[1]), aes(Param, min_time)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept=0.05, lty=2) + 
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('a. {Parameter}')) +
  ylim(10,50) +
  labs(x=c("sigma"), y= "Minimum number of samples required") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))
p2 <- ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[2]), aes(Param, min_time)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 0.404, lty=2) +
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('b. {Parameter}')) +
  ylim(10,50) +
  labs(x="phi", y= "") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))
p3 <- ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[3]), aes(Param, min_time))  +
  geom_point() +
  geom_line() + 
  geom_vline(xintercept = 0.81, lty=2) +
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('c. {Parameter}')) +
  ylim(10,50) +
  labs(x="shift size", y= "") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))

# Remove the y-axis for the 2nd plot - p2 then merge 2 plots
cowplot::plot_grid(p1 + theme(plot.margin = unit(c(0.15, 0.1, 0.15, 0.15), "cm")), 
                   p2 + 
                     theme(plot.margin = unit(c(0, 0.1, 0, 0), "cm"),
                           axis.text.y = element_blank(),
                           axis.line.y = element_blank(),
                           axis.title.y= element_blank(),
                           axis.ticks.y= element_blank()),
                   p3 + 
                     theme(plot.margin = unit(c(0, 0.1, 0, 0), "cm"),
                           axis.text.y = element_blank(),
                           axis.line.y = element_blank(),
                           axis.title.y= element_blank(),
                           axis.ticks.y= element_blank()),
                   nrow = 1,
                   rel_widths = c(1.2, 1,1),
                   align = 'h', axis = 'tb')

```

## Iterative sampling

```{r simulation_min_time_iterative,message=F,warning=F,echo=F,eval=F,cache=T}

source('R/sampling_iterative.R')
source('R/build_time_series.R')
source('R/updated_param_sensitivity.R')
source('R/calc_min_time2.R')
require(tidyverse)

# Sigma results
sigma_vec = seq(0.01,0.4,0.025)
min_time  = rep(0,times=length(sigma_vec))
sigma_output <- data.frame(sigma_vec,min_time)

for (param_num in 1:nrow(sigma_output)){
  sigma_output$min_time[param_num] <- calc_min_time('sigma',sigma_output$sigma_vec[param_num],'iterative')
  #print(sigma_output)
}

write.csv(x = sigma_output,file = 'Output/theoretical_results_iterative_sigma.csv',quote=FALSE)

# Phi results

phi_vec = seq(-0.8,0.8,0.05)
min_time  = rep(0,times=length(phi_vec))
phi_output <- data.frame(phi_vec,min_time)

for (param_num in 1:nrow(phi_output)){
  phi_output$min_time[param_num] <- calc_min_time('phi',phi_output$phi_vec[param_num],'iterative')
  #print(phi_output)
}

write.csv(x = phi_output,file = 'Output/theoretical_results_iterative_phi.csv',quote=FALSE)

# Shift size

shift_size_vec = rev(seq(0.1,0.9,0.05))
min_time  = rep(0,times=length(shift_size_vec))
shift_size_output <- data.frame(shift_size_vec,min_time)

for (param_num in 1:nrow(shift_size_output)){
  shift_size_output$min_time[param_num] <- calc_min_time('shift_size',shift_size_output$shift_size_vec[param_num],'iterative')
  #print(shift_size_output)
}

write.csv(x = shift_size_output,file = 'Output/theoretical_results_iterative_shift_size.csv',quote=FALSE)

```

```{r simulation_min_time_plot_iterative, fig.width = 7.2, fig.height= 3, fig.cap='Minimum number of samples required for 0.8 statistical power given different levels of (a) temporal variability ($\\sigma$), (b) temporal autocorrelation ($\\phi$), and (c) shift size. Iterative sampling was used with the default parameters are $\\sigma = 0.53$, $\\phi = 0.404$, and a shift size $=0.81$ to match the case study. The exact timing of the true changepoint varied for each simulation between time steps 30 and 70. Each vertical line indicates the respective parameter calculated from the case study time series. \\label{fig:simulation_min_time}',message=F,warning=F,echo=F,eval=T}

par(mfrow=c(1,3))

sigma <- read.csv('Output/theoretical_results_iterative_sigma.csv',header=TRUE)

phi <- read.csv('Output/theoretical_results_iterative_phi.csv',header=TRUE)

shift_size <- read.csv('Output/theoretical_results_iterative_shift_size.csv',header=TRUE)

# Melt the 3 dataset to be able to use facet_wrap()
colnames(sigma) <- colnames(phi) <- colnames(shift_size) <- c("X", "Param", "min_time")
simulation_all <- 
  rbind(cbind("Parameter" = rep("Standard deviation", nrow(sigma)), sigma),
        cbind("Parameter" = rep("Autocorrelation at lag-1", nrow(phi)), phi),
        cbind("Parameter" = rep("Normalized amplitude of shift", nrow(shift_size)), shift_size))
# param_vr <- data.frame("Parameter" = unique(simulation_all$Parameter),
#                        "value" = c(ar_vr, cv_vr, shift_vr))

p1 <- 
  ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[1]), aes(Param, min_time)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept=0.05, lty=2) + 
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('a. {Parameter}')) +
  ylim(10,50) +
  labs(x=c("sigma"), y= "Minimum number of samples required") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))
p2 <- ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[2]), aes(Param, min_time)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 0.404, lty=2) +
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('b. {Parameter}')) +
  ylim(10,50) +
  labs(x="phi", y= "") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))
p3 <- ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[3]), aes(Param, min_time))  +
  geom_point() +
  geom_line() + 
  geom_vline(xintercept = 0.81, lty=2) +
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('c. {Parameter}')) +
  ylim(10,50) +
  labs(x="shift size", y= "") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))

# Remove the y-axis for the 2nd plot - p2 then merge 2 plots
cowplot::plot_grid(p1 + theme(plot.margin = unit(c(0.15, 0.1, 0.15, 0.15), "cm")), 
                   p2 + 
                     theme(plot.margin = unit(c(0, 0.1, 0, 0), "cm"),
                           axis.text.y = element_blank(),
                           axis.line.y = element_blank(),
                           axis.title.y= element_blank(),
                           axis.ticks.y= element_blank()),
                   p3 + 
                     theme(plot.margin = unit(c(0, 0.1, 0, 0), "cm"),
                           axis.text.y = element_blank(),
                           axis.line.y = element_blank(),
                           axis.title.y= element_blank(),
                           axis.ticks.y= element_blank()),
                   nrow = 1,
                   rel_widths = c(1.2, 1,1),
                   align = 'h', axis = 'tb')

```

# Case study

```{r case study read data output, include=FALSE}
myoutput <- read.table(paste0(getwd(),"/Output/output_changepoint_varese.txt"))
myoutput$method_f <- factor(myoutput$method, levels=c('Random', 'Regular', 'Iterative'))

```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 7.2, fig.height= 8}
library(ggplot2)
library(gridExtra)
p1 <- ggplot(myoutput[myoutput$target_cpt==1,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="a. First changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
p2 <- ggplot(myoutput[myoutput$target_cpt==2,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="b. Second changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
p3 <- ggplot(myoutput[myoutput$target_cpt==3,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="c. Third changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
grid.arrange(p1,p2,p3,nrow=3)
```

\clearpage

## References