### 2020-03-31 R/supp_material.Rmd
first edits of the day to supp material

---
bibliography: Whitebib.bib
csl: ecology-letters.csl
editor_options:
  chunk_output_type: console
fontsize: 12pt
geometry: margin=1in
header-includes: \usepackage{float} \usepackage{lineno} \usepackage{setspace}\doublespacing
  \usepackage[round]{natbib} \bibpunct[; ]{(}{)}{,}{a}{}{,} \usepackage{color} \usepackage{totcount}
  \newtotcounter{citenum} \def\oldcite{} \let\oldcite=\bibcite \def\bibcite{\stepcounter{citenum}\oldcite}
  \renewcommand{\thepage}{S\arabic{page}}
  \renewcommand{\thesection}{Appendix S\arabic{section}} \renewcommand{\thetable}{S\arabic{table}}
  \renewcommand{\thefigure}{S\arabic{figure}}
  \usepackage{amsmath}
  \renewcommand{\theequation}{S\arabic{equation}} \floatplacement{figure}{H}
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = dirname(getwd()))
```

\begin{center}
\textbf{\Large Supplementary Material for: Sampling requirements and approaches to detect ecosystem shifts}
\vspace{7 mm}

\textsc{Rosalie Bruel$^{1}$ and Easton R. White$^{2}$}
\vspace{5 mm}

\normalsize{\indent $^{1}$Rubenstein Ecosystem Science Laboratory, University of Vermont, VT 05405, USA \\ $^{2}$Department of Biology, University of Vermont, VT 05405, USA}
\end{center}

\vspace{2 mm}

\tableofcontents

\vspace{1cm}

Data and code for all the figures can be found at (https://github.com/rosalieb/temporal-sampling.git). All the analyses were run in R (citation).

\clearpage

# Additional methods

We first detect the ``true" changepoint of a full length time series with the function _e.divisive_ in the R package _ecp_ [@James2019]. We focus on the changepoint with the largest magnitude, altough this package allows detection of further changepoints as well (Fig. \ref{fig:case_study_other_changepoints}). We then subsample the full time series [@White2020] with different numbers of subsamples and different sampling approaches. For each sampling approach, we wrote a custom function:

- sample_random(): sample first and last sample, in addition of n-2 random samples in between (n = maximum number of sample, chosen by the user),  
- sample_regular(): sample first and last sample, in addition of n-2 evenly spaced samples in between (n = maximum number of samples, chosen by the user), and  
- sample_iterative(): sample first and last sample as well as 3 evenly spaced samples in between (3 is the default but can be modified), in order to initiate the changepoint detection. Upon detection of the changepoint on this 5 samples time series, a new sample is added between the changepoint and the previous sample, to narrow down the real changepoint. Detection of changepoint and addition of sample is repeated, until it was narrowed down to two consecutive samples, or n (maximum number of samples, chosen by the user) was reached, whichever comes first.

The subsampled time series are then compared to the full time series to assess the effectiveness of each subsampling approach.

Each function works with two types of input: vector and matrix. If the input is a matrix, the user must change the default argument *is_vector* to *FALSE*. The matrix is transformed to independent vectors using Detrented Correspondance Analysis [@hill_detrended_1980], and a single axis on which changepoint analysis is run is chosen with the argument *DCA_axis* (default to 1, for first component).

# Parameter sensitivity

```{r parameter_sensitivity,message=F,warning=F,echo=F,eval=F,cache=F}
# Build 1x3 plot with sigma vs phi with color equal to distance to changepoint, power, or min time. The three panel plots would represent different shift size values

source('R/sampling_regular.R')
source('R/sampling_random.R')
source('R/build_time_series.R')
source('R/calc_min_time2.R')
require(tidyverse)

number_sims <- 10
# Find minimum time for different parameter combos

sigma_vec = seq(0.01,0.3,0.02)
phi_vec = seq(-0.8,0.8,0.1)
shift_size_vec = c(0.1,0.4,0.8)

output_theory_approach <- data.frame(expand.grid(sigma_vec,phi_vec,shift_size_vec))
names(output_theory_approach) <- c('sigma_vec', 'phi_vec', 'shift_size_vec')
output_theory_approach$power = 0

#params <- tibble(shift_time = sample(30:70,max(simulation_number),replace = T))

shift_time = sample(30:70,number_sims,replace=T) # Keep shift times constant for each param combo

# Run through params
for (i in 1:nrow(output_theory_approach)){

# Build XX number of time series
params <- list(sigma=rep(output_theory_approach$sigma_vec[i],number_sims),phi=rep(output_theory_approach$phi_vec[i],number_sims),shift_size=rep(output_theory_approach$shift_size_vec[i],number_sims),shift_time = shift_time)

time_series <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) sample_regular(df,20)$changepoint)

power <- sum(abs(time_series-shift_time)<5)/number_sims

#print(i)
output_theory_approach$power[i] <- power
}

write.csv(x = output_theory_approach,file = 'Output/parameter_sensitivity.csv',quote = FALSE )

require(ggplot2)
require(viridis)
ggplot(aes(x=phi_vec,y=sigma_vec,fill=power),data=output_theory_approach) + geom_tile() +
   scale_fill_viridis(discrete=FALSE,begin = 0.1,end=0.9) +
  facet_wrap(~shift_size_vec) +
   labs(fill='Statistical power') +
   xlab("Lag-1 autocorrelation (phi)") +
   ylab("Standard deviation (sigma)") +
   theme_classic(base_size = 14, base_family = "")

```

```{r parameter_sensitivity_plot,echo=F,message=F,warnings=F,fig.height=4,fig.cap='Regular sampling statistical power (fraction of 100 simulations which detecting a changepoint within five time points of the true changepoint) for different levels of standard deviation, lag-1 autocorrelation, and shift size.\\label{fig:parameter_sensitivity_plot}'}
output_theory_approach <- read.csv(file = 'Output/parameter_sensitivity.csv')

require(ggplot2)
require(viridis)
ggplot(aes(x=phi_vec,y=sigma_vec,fill=power),data=output_theory_approach) + geom_tile() +
   scale_fill_viridis(discrete=FALSE,begin = 0.1,end=0.9) +
  facet_wrap(~shift_size_vec) + 
   theme_classic(base_size = 14, base_family = "")+
  theme(plot.subtitle = element_text(hjust = 0.5)) +
   labs(fill='Power',subtitle = 'Shift size') +
   xlab("Lag-1 autocorrelation (phi)") +
   ylab("Standard deviation (sigma)")
```

# Minimum time for other sampling approaches

## Random sampling

```{r simulation_min_time_random,message=F,warning=F,echo=F,eval=F,cache=T}

source('R/sampling_random.R')
source('R/build_time_series.R')
source('R/updated_param_sensitivity.R')
source('R/calc_min_time2.R')
require(tidyverse)

# Sigma results
sigma_vec = seq(0.01,0.4,0.025)
min_time  = rep(0,times=length(sigma_vec))
sigma_output <- data.frame(sigma_vec,min_time)

for (param_num in 1:nrow(sigma_output)){
  sigma_output$min_time[param_num] <- calc_min_time('sigma',sigma_output$sigma_vec[param_num],'random')
  #print(sigma_output)
}

write.csv(x = sigma_output,file = 'Output/theoretical_results_random_sigma.csv',quote=FALSE)

# Phi results

phi_vec = seq(-0.8,0.8,0.05)
min_time  = rep(0,times=length(phi_vec))
phi_output <- data.frame(phi_vec,min_time)

for (param_num in 1:nrow(phi_output)){
  phi_output$min_time[param_num] <- calc_min_time('phi',phi_output$phi_vec[param_num],'random')
  #print(phi_output)
}

write.csv(x = phi_output,file = 'Output/theoretical_results_random_phi.csv',quote=FALSE)

# Shift size

shift_size_vec = rev(seq(0.1,0.9,0.05))
min_time  = rep(0,times=length(shift_size_vec))
shift_size_output <- data.frame(shift_size_vec,min_time)

for (param_num in 1:nrow(shift_size_output)){
  shift_size_output$min_time[param_num] <- calc_min_time('shift_size',shift_size_output$shift_size_vec[param_num],'random')
  #print(shift_size_output)
}

write.csv(x = shift_size_output,file = 'Output/theoretical_results_random_shift_size.csv',quote=FALSE)

```

```{r simulation_min_time_plot_random, fig.width = 7.2, fig.height= 3, fig.cap='Minimum number of samples required for 0.8 statistical power given different levels of (a) temporal variability ($\\sigma$), (b) temporal autocorrelation ($\\phi$), and (c) shift size. Random sampling was used with the default parameters are $\\sigma = 0.53$, $\\phi = 0.404$, and a shift size $=0.81$ to match the case study. The exact timing of the true changepoint varied for each simulation between time steps 30 and 70. Each vertical line indicates the respective parameter calculated from the case study time series. \\label{fig:simulation_min_time_random}',message=F,warning=F,echo=F,eval=T}
require(ggplot2)
require(dplyr)
require(stickylabeller)
par(mfrow=c(1,3))

sigma <- read.csv('Output/theoretical_results_random_sigma.csv',header=TRUE)

phi <- read.csv('Output/theoretical_results_random_phi.csv',header=TRUE)

shift_size <- read.csv('Output/theoretical_results_random_shift_size.csv',header=TRUE)

# Melt the 3 dataset to be able to use facet_wrap()
colnames(sigma) <- colnames(phi) <- colnames(shift_size) <- c("X", "Param", "min_time")
simulation_all <- 
  rbind(cbind("Parameter" = rep("Standard deviation", nrow(sigma)), sigma),
        cbind("Parameter" = rep("Autocorrelation at lag-1", nrow(phi)), phi),
        cbind("Parameter" = rep("Normalized amplitude of shift", nrow(shift_size)), shift_size))
# param_vr <- data.frame("Parameter" = unique(simulation_all$Parameter),
#                        "value" = c(ar_vr, cv_vr, shift_vr))

p1 <- 
  ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[1]), aes(Param, min_time)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept=0.05, lty=2) + 
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('a. {Parameter}')) +
  ylim(10,52) +
  labs(x=c("sigma"), y= "Minimum number of samples required") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))
p2 <- ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[2]), aes(Param, min_time)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 0.404, lty=2) +
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('b. {Parameter}')) +
  ylim(10,52) +
  labs(x="phi", y= "") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))
p3 <- ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[3]), aes(Param, min_time))  +
  geom_point() +
  geom_line() + 
  geom_vline(xintercept = 0.81, lty=2) +
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('c. {Parameter}')) +
  ylim(10,52) +
  labs(x="shift size", y= "") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))

# Remove the y-axis for the 2nd plot - p2 then merge 2 plots
cowplot::plot_grid(p1 + theme(plot.margin = unit(c(0.15, 0.1, 0.15, 0.15), "cm")), 
                   p2 + 
                     theme(plot.margin = unit(c(0, 0.1, 0, 0), "cm"),
                           axis.text.y = element_blank(),
                           axis.line.y = element_blank(),
                           axis.title.y= element_blank(),
                           axis.ticks.y= element_blank()),
                   p3 + 
                     theme(plot.margin = unit(c(0, 0.1, 0, 0), "cm"),
                           axis.text.y = element_blank(),
                           axis.line.y = element_blank(),
                           axis.title.y= element_blank(),
                           axis.ticks.y= element_blank()),
                   nrow = 1,
                   rel_widths = c(1.2, 1,1),
                   align = 'h', axis = 'tb')

```

## Iterative sampling

```{r simulation_min_time_iterative,message=F,warning=F,echo=F,eval=F,cache=T}

source('R/sampling_iterative.R')
source('R/build_time_series.R')
source('R/updated_param_sensitivity.R')
source('R/calc_min_time2.R')
require(tidyverse)

# Sigma results
sigma_vec = seq(0.01,0.4,0.025)
min_time  = rep(0,times=length(sigma_vec))
sigma_output <- data.frame(sigma_vec,min_time)

for (param_num in 1:nrow(sigma_output)){
  sigma_output$min_time[param_num] <- calc_min_time('sigma',sigma_output$sigma_vec[param_num],'iterative')
  #print(sigma_output)
}

write.csv(x = sigma_output,file = 'Output/theoretical_results_iterative_sigma.csv',quote=FALSE)

# Phi results

phi_vec = seq(-0.8,0.8,0.05)
min_time  = rep(0,times=length(phi_vec))
phi_output <- data.frame(phi_vec,min_time)

for (param_num in 1:nrow(phi_output)){
  phi_output$min_time[param_num] <- calc_min_time('phi',phi_output$phi_vec[param_num],'iterative')
  #print(phi_output)
}

write.csv(x = phi_output,file = 'Output/theoretical_results_iterative_phi.csv',quote=FALSE)

# Shift size

shift_size_vec = rev(seq(0.1,0.9,0.05))
min_time  = rep(0,times=length(shift_size_vec))
shift_size_output <- data.frame(shift_size_vec,min_time)

for (param_num in 1:nrow(shift_size_output)){
  shift_size_output$min_time[param_num] <- calc_min_time('shift_size',shift_size_output$shift_size_vec[param_num],'iterative')
  #print(shift_size_output)
}

write.csv(x = shift_size_output,file = 'Output/theoretical_results_iterative_shift_size.csv',quote=FALSE)

```

```{r simulation_min_time_plot_iterative, fig.width = 7.2, fig.height= 3, fig.cap='Minimum number of samples required for 0.8 statistical power given different levels of (a) temporal variability ($\\sigma$), (b) temporal autocorrelation ($\\phi$), and (c) shift size. Iterative sampling was used with the default parameters are $\\sigma = 0.53$, $\\phi = 0.404$, and a shift size $=0.81$ to match the case study. The exact timing of the true changepoint varied for each simulation between time steps 30 and 70. Each vertical line indicates the respective parameter calculated from the case study time series. \\label{fig:simulation_min_time}',message=F,warning=F,echo=F,eval=T}

par(mfrow=c(1,3))

sigma <- read.csv('Output/theoretical_results_iterative_sigma.csv',header=TRUE)

phi <- read.csv('Output/theoretical_results_iterative_phi.csv',header=TRUE)

shift_size <- read.csv('Output/theoretical_results_iterative_shift_size.csv',header=TRUE)

# Melt the 3 dataset to be able to use facet_wrap()
colnames(sigma) <- colnames(phi) <- colnames(shift_size) <- c("X", "Param", "min_time")
simulation_all <- 
  rbind(cbind("Parameter" = rep("Standard deviation", nrow(sigma)), sigma),
        cbind("Parameter" = rep("Autocorrelation at lag-1", nrow(phi)), phi),
        cbind("Parameter" = rep("Normalized amplitude of shift", nrow(shift_size)), shift_size))
# param_vr <- data.frame("Parameter" = unique(simulation_all$Parameter),
#                        "value" = c(ar_vr, cv_vr, shift_vr))

p1 <- 
  ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[1]), aes(Param, min_time)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept=0.05, lty=2) + 
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('a. {Parameter}')) +
  ylim(10,50) +
  labs(x=c("sigma"), y= "Minimum number of samples required") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))
p2 <- ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[2]), aes(Param, min_time)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 0.404, lty=2) +
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('b. {Parameter}')) +
  ylim(10,50) +
  labs(x="phi", y= "") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))
p3 <- ggplot(simulation_all %>% filter(Parameter == unique(simulation_all$Parameter)[3]), aes(Param, min_time))  +
  geom_point() +
  geom_line() + 
  geom_vline(xintercept = 0.81, lty=2) +
  facet_wrap(~Parameter, scales="free_x", 
             labeller = label_glue('c. {Parameter}')) +
  ylim(10,50) +
  labs(x="shift size", y= "") + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0))

# Remove the y-axis for the 2nd plot - p2 then merge 2 plots
cowplot::plot_grid(p1 + theme(plot.margin = unit(c(0.15, 0.1, 0.15, 0.15), "cm")), 
                   p2 + 
                     theme(plot.margin = unit(c(0, 0.1, 0, 0), "cm"),
                           axis.text.y = element_blank(),
                           axis.line.y = element_blank(),
                           axis.title.y= element_blank(),
                           axis.ticks.y= element_blank()),
                   p3 + 
                     theme(plot.margin = unit(c(0, 0.1, 0, 0), "cm"),
                           axis.text.y = element_blank(),
                           axis.line.y = element_blank(),
                           axis.title.y= element_blank(),
                           axis.ticks.y= element_blank()),
                   nrow = 1,
                   rel_widths = c(1.2, 1,1),
                   align = 'h', axis = 'tb')

```

# Case study

```{r case study read data output, include=FALSE}
myoutput <- read.table(paste0(getwd(),"/Output/output_changepoint_varese.txt"))
myoutput$method_f <- factor(myoutput$method, levels=c('Random', 'Regular', 'Iterative'))

```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 7.2, fig.height= 8,fig.cap='Distance to true (a) first changepoint (largest magnitude), (b) second changepoint, and (c) third changepoint for total number of samples analyzed, following random sampling, regular sampling, and iterative sampling. Total number of samples was set between 5 and 30, out of the 74 initial time series.  \\label{fig:case_study_other_changepoints}'}
library(ggplot2)
library(gridExtra)
p1 <- ggplot(myoutput[myoutput$target_cpt==1,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="a. First changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
p2 <- ggplot(myoutput[myoutput$target_cpt==2,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="b. Second changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
p3 <- ggplot(myoutput[myoutput$target_cpt==3,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="c. Third changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
grid.arrange(p1,p2,p3,nrow=3)
```

## Case study: biomass

```{r echo=FALSE, eval=T,message=FALSE, warning=FALSE, fig.width = 7.2, fig.height= 3,fig.cap='Distance to real changepoint in biomass per total number of samples analyzed, following (a) random sampling, (b) regular sampling, and (c) iterative sampling. Total number of samples was set between 5 and 30, out of the 74 initial time series.  \\label{fig:case_study_number_sample}'}
#investigating species average biomass
vr_count <- rowSums(vr[,-1])
myoutput_count <- NULL
j=3
for (i in 5:30) {
  vrsub <- sample_regular(vr_count, i, input_vector = T, messages = F)
  myoutput_count <- c(myoutput_count, vrsub$method, vrsub$final_n, vrsub$changepoint_all$order.found[j], length(vrsub$changepoint_all$estimates)-2)
  vrsub <- sample_random(vr_count, i, input_vector = T, messages = F)
  myoutput_count <- c(myoutput_count, vrsub$method, vrsub$final_n, vrsub$changepoint_all$order.found[j], length(vrsub$changepoint_all$estimates)-2)
  vrsub <- sample_iterative(vr_count, i, input_vector = T, messages = F)
  myoutput_count <- c(myoutput_count, vrsub$method, vrsub$final_n, vrsub$changepoint_all$order.found[j], length(vrsub$changepoint_all$estimates)-2)
  vrsub <- sample_iterative(vr_count, i, input_vector = T, messages = F, n2 = 6)
  myoutput_count <- c(myoutput_count, vrsub$method, vrsub$final_n, vrsub$changepoint_all$order.found[j], length(vrsub$changepoint_all$estimates)-2)
  vrsub <- sample_iterative(vr_count, i, input_vector = T, messages = F, n2 = 7)
  myoutput_count <- c(myoutput_count, vrsub$method, vrsub$final_n, vrsub$changepoint_all$order.found[j], length(vrsub$changepoint_all$estimates)-2)
}

myoutput <- read.table(paste0(getwd(),"/Output/output_changepoint_varese_biomass.txt"))
myoutput$method <- factor(myoutput$method, levels=c('Random', 'Regular', 'Iterative'))
myoutput$diff_real <- abs(myoutput$diff_real)

# Plot
p1 <-
  ggplot(myoutput[myoutput$target_cpt==1 & myoutput$n2_init == 5,], aes(final_n,diff_real)) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method,
             labeller = label_glue('{.l}. {method} sampling ')) + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="transparent", fill="transparent"),
        strip.text.x = element_text(angle = 0, hjust = 0)) +
  labs(x="Total number of samples analyzed", y="Distance to real changepoint")

p1

```

(Fig. \ref{fig:case_study_number_sample_change_biomass})

\clearpage

## References